---
date: 2017-06-21T20:01:45-04:00
subtitle: "Not too many."
type: "post"
tags: []
title: "How many profiles do Free users use?"
---

## Introduction
There is a big focus on simplifying the product for Buffer users on free plans in 2017 in order to create a more intuitive experience and encourage more upgrades.  

The purpose of this analysis is to get a better understanding of how users currently interact with the free plan, in order to better inform decisions around future plan limits. Specifically, we would love to learn about how many profiles users connect and actually schedule updates for. 

More data questions and context can be found in [**this Paper**](https://paper.dropbox.com/doc/Free-Plan-data-KJNJLDW3KtfMdZK4Z6DRS), and potential changes to the Free plan can be found in [**this Paper**](https://paper.dropbox.com/doc/Free-plan-changes-DVheYkvNs3yiE9Nhb9XRF).

## Methodology
To answer this question, we'd like to only consider users that are currently _active_, and analyze their usage _during the period in which they were active in Buffer_. In this case, _active_ is defined as having scheduled at least one update with Buffer in the previous 28 days.

We'll gather all of the currently active users that are on free plans. Then, we will gather data on the number of profiles they have connected and the number of updates that they have scheduled _for each profile_.

Once we have this data, we will gather some summary statistics on the number of active profiles for each active user, and visualize the distribution of the number of profiles in use. It is unlikely to be normally distributed (bell shaped), so visualizing this breakdown would be useful to understanding the potential impact that changing the limits might have.

## Data collection
Let's start by getting the Free users that have scheduled at least one update in the past 28 days. We'll use the following SQL query to grab them.

```{r include = FALSE}
library(buffer); library(dplyr); library(ggplot2)
```

```{r include = FALSE}
con <- redshift_connect()
```

```{sql connection=con, eval = FALSE}
with user_facts as (
  select
    up.user_id
    , max(up.created_at) as last_update_created_at
  from transformed_updates as up
  inner join users
    on users.user_id = up.user_id
  where up.status != 'service'
  and users.billing_plan = 'individual'
  group by up.user_id
)
select
  user_id
  , last_update_created_at
from user_facts
where last_update_created_at > (current_date - 29)
```

```{r include = F}
# Save users
# save(users, file = "user_profiles.Rda")

# Load users data
load("user_profiles.Rda")
```

There are around 250 thousand users in this dataset! Now we need to get the number of updates they have scheduled, and the number of profiles that they have scheduled updates for.

```{sql connection=con, eval = FALSE}
select
  up.user_id
  , count(distinct up.id) as updates_count
  , count(distinct up.profile_id) as profile_count
from transformed_updates as up
left join users
  on up.user_id = users.user_id
where users.billing_plan = 'individual'
and up.created_at > (current_date - 29)
group by up.user_id
```

```{r include = F}
# Save updates
# save(updates, file = "user_profile_updates.Rda")

# Load users data
load("user_profile_updates.Rda")
```

Awesome! Now let's join the `updates` and `users` dataframes so that we'll have the number of updates and active profiles for each user. We'll use `dplyr`'s `left_join` function and replace `NA` values with 0.

```{r}
# Join updates and users
users <- users %>%
  left_join(updates, by = 'user_id')

# Replace NAs with 0s
users[is.na(users)] <- 0
```

## Exploratory analysis
Now that we have the updates and profile counts for each active user, we can compute some quick summary statistics for our population.

```{r}
# Get summary stats on the number of profiles
summary(users$profile_count)
```

Alright, here is the breakdown:

 - Around 50% of active Free plan users scheduled updates for only 1 profile in the past 28 days.
 - The average number of profiles that active Free plan users used in the past 28 days is 1.85.
 - Around 75% of active Free plan users scheduled updates for only 2 or less profiles.
 - Someone scheduled updates for 194 profiles in the past 28 days!
 