---
date: 2017-06-17T20:01:45-04:00
subtitle: "Not too many."
type: "post"
tags: []
title: "How many profiles do Free users use?"
---

## Motivation 
The purpose of this analysis is to get a better understanding of how users on the Individual plan use Buffer, in order to better inform decisions around the plan limits. Specifically, we would love to learn about how many profiles Free users actually schedule updates for. 

More data questions and context can be found in [**this Paper**](https://paper.dropbox.com/doc/Free-Plan-data-KJNJLDW3KtfMdZK4Z6DRS), and potential changes to the Free plan can be found in [**this Paper**](https://paper.dropbox.com/doc/Free-plan-changes-DVheYkvNs3yiE9Nhb9XRF).

## Methodology
In order to answer this question reliably with data, we would like to only consider users that are currently _active_, and analyze their updates _during the period in which they were active in Buffer_. In this case, _active_ is defined as having scheduled at least one update with Buffer in the past 28 days.

To answer the question of how many profiles Free users actually use, we'll first gather all of the currently active Free users. Then, we will gather data on the number of profiles they have, and the number of updates that they have scheduled _for each profile_.

Once we have this data, we will gather some summary statistics on the number of active profiles for each active user, and visualize the distribution of the number of profiles in use. It is unlikely to be normally distributed (bell shaped), so visualizing this breakdown would be useful to understanding the potential impact that changing the limits might have.

## Data collection
Let's start by getting the Free users that have scheduled at least one update in the past 28 days. We'll use the following SQL query to grab them.

```{r include = FALSE}
library(buffer); library(dplyr); library(ggplot2)
```

```{r include = FALSE}
con <- redshift_connect()
```

```{sql connection=con, eval = FALSE}
with user_facts as (
  select
    up.user_id
    , max(up.created_at) as last_update_created_at
  from transformed_updates as up
  inner join users
    on users.user_id = up.user_id
  where up.status != 'service'
  and users.billing_plan = 'individual'
  group by up.user_id
)
select
  user_id
  , last_update_created_at
from user_facts
where last_update_created_at > (current_date - 29)
```

```{r include = F}
# Save users
# save(users, file = "user_profiles.Rda")

# Load users data
load("user_profiles.Rda")
```

```{r}
head(users)
```

