---
date: 2017-06-21T20:01:45-04:00
subtitle: "Not too many."
bigimg: [{src: "/blog/profile_analysis_files/figure-html/unnamed-chunk-10-1.png"}]
type: "post"
tags: []
title: "How many profiles do Free users use?"
---

## Introduction
There is a big focus on simplifying the product for Buffer users on free plans in 2017 in order to create a more intuitive experience and encourage more upgrades.  

The purpose of this analysis is to get a better understanding of how users currently interact with the free plan, in order to better inform decisions around future plan limits. Specifically, we would love to learn about how many profiles users connect and actually schedule updates for. 

More data questions and context can be found in [**this Paper**](https://paper.dropbox.com/doc/Free-Plan-data-KJNJLDW3KtfMdZK4Z6DRS), and potential changes to the Free plan can be found in [**this Paper**](https://paper.dropbox.com/doc/Free-plan-changes-DVheYkvNs3yiE9Nhb9XRF).

## Methodology
To answer this question, we'd like to only consider users that are currently _active_, and analyze their usage _during the period in which they were active in Buffer_. In this case, _active_ is defined as having scheduled at least one update with Buffer in the previous 28 days.

We'll gather all of the currently active users that are on free plans. Then, we will gather data on the number of profiles they have connected and the number of updates that they have scheduled _for each profile_.

Once we have this data, we will gather some summary statistics on the number of active profiles for each active user, and visualize the distribution of the number of profiles in use. It is unlikely to be normally distributed (bell shaped), so visualizing this breakdown would be useful to understanding the potential impact that changing the limits might have.

## Data collection
Let's start by getting the Free users that have scheduled at least one update in the past 28 days. We'll use the following SQL query to grab them.

```{r include = FALSE}
library(buffer); library(dplyr); library(ggplot2)
```

```{r include = FALSE}
con <- redshift_connect()
```

```{sql connection=con, eval=FALSE}
with user_facts as (
  select
    up.user_id
    , max(up.created_at) as last_update_created_at
  from transformed_updates as up
  inner join users
    on users.user_id = up.user_id
  where up.status != 'service'
  and users.billing_plan = 'individual'
  group by up.user_id
)
select
  user_id
  , last_update_created_at
from user_facts
where last_update_created_at > (current_date - 29)
```

```{r include = F}
# Save users
# save(users, file = "user_profiles.Rda")

# Load users data
load("user_profiles.Rda")
```

There are around 250 thousand users in this dataset! Now we need to get the number of updates they have scheduled, and the number of profiles that they have scheduled updates for.

```{sql connection=con, eval=FALSE}
select
  up.user_id
  , count(distinct up.id) as updates_count
  , count(distinct up.profile_id) as profile_count
from transformed_updates as up
left join users
  on up.user_id = users.user_id
where users.billing_plan = 'individual'
and up.status != 'service'
and up.created_at > (current_date - 29)
group by up.user_id
```

```{r include = F}
# Save updates
# save(updates, file = "user_profile_updates.Rda")

# Load users data
load("user_profile_updates.Rda")
```

Awesome! Now let's join the `updates` and `users` dataframes so that we'll have the number of updates and active profiles for each user. We'll use `dplyr`'s `left_join` function and replace `NA` values with 0.

```{r}
# Join updates and users
users <- users %>%
  left_join(updates, by = 'user_id')

# Replace NAs with 0s
users[is.na(users)] <- 0
```

## Exploratory analysis
Now that we have the updates and profile counts for each active user, we can compute some quick summary statistics for our population.

```{r}
# Get summary stats on the number of profiles
summary(users$profile_count)
```

Here is the breakdown:

 - Around 50% of active Free plan users scheduled updates for only 1 profile in the past 28 days.
 - The average number of profiles that active Free plan users used in the past 28 days is around 1.9.
 - Around 75% of active Free plan users scheduled updates for only 2 or less profiles.
 - Someone scheduled updates for 194 profiles in the past 28 days!
 
The maximum of 194 brings up a good point about our sample: these are only users that are _currently_ on individual plans. This means that they could have been on a paid plan in the past 28 days. I don't think it's a huge portion, or very significant for this particular analysis, but it is something to keep in mind.

Let's look at the distribution of profiles for active free users.

```{r include = FALSE}
by_profile <- users %>%
  group_by(profile_count) %>%
  summarise(users = n_distinct(user_id)) %>%
  mutate(proportion = users / sum(users))
```

```{r echo = FALSE}
options(scipen=10000)

ggplot(by_profile, aes(x = as.factor(profile_count), y = users)) +
  geom_bar(stat = 'identity') +
  coord_cartesian(xlim = c(1,10)) +
  geom_text(aes(label = users), position = position_dodge(width=0.9), vjust=-0.25) +
  labs(x = 'Number of Active Profiles', y = 'Users')
```

Wow, that's interesting! It looks like most currently active Free users have only scheduled updates to one profile in the past 28 days. 

One possibility is that there are very new users that have only gotten the chance to schedule updates to one or two profiles. Let's see what would happen if we only looked at users that have sent at least 5 updates in the past 28 days.

```{r echo = FALSE}
# Group by number of profiles
by_profile <- users %>%
  filter(updates_count >= 5) %>%
  group_by(profile_count) %>%
  summarise(users = n_distinct(user_id)) %>%
  mutate(proportion = users / sum(users))

# Build bar plot
ggplot(by_profile, aes(x = as.factor(profile_count), y = users)) +
  geom_bar(stat = 'identity') +
  coord_cartesian(xlim = c(1,10)) +
  geom_text(aes(label = users), position = position_dodge(width=0.9), vjust=-0.25) +
  labs(x = 'Number of Active Profiles', y = 'Users')
```

We can see here that the proportion of active users that only schedule to a single profile changes quite a bit, but it is still high at 46.5%. 

Let's look at the CDF of the profiles distribution.

```{r echo = FALSE}
ggplot(users, aes(x = profile_count)) + 
  stat_ecdf() +
  coord_cartesian(xlim = c(0, 10)) + 
  scale_x_continuous(breaks = seq(0, 10, 2)) +
  scale_y_continuous(breaks = seq(0, 1, 0.2)) +
  labs(x = 'Number of Active Profiles', y = '' , title = 'Profile Count CDF')
```

Interesting! This graph tells us the following:

 - Around 56% of users scheduled updates to 1 profile in the past 28 days.
 - Around 76% of users scheduled updates to 2 or less profiles in the past 28 days.
 - Around 91% of users scheduled updates to 3 or less profiles in the past 28 days.
 - Around 98% of users scheduled updates to 4 or less profiles in the past 28 days.
 
## Conclusions
If we were to go with the "less limited" Free plan and only allow users to connect up to 3 profiles, the data suggests that somewhere around 10% of currently active users would be affected. This equates to around 25k users.

If we were to go with the "more limited" Free plan and only allow 2 profiles, the data suggests that somewhere around 25% of currently active users would be affected! This equates to around 63k users.

I think it's important to note that this is just a snapshot of _currently_ active users that are _currently_ on Individual plans. 

There are still important things that we **don't** know as of yet. We don't know how these users with 3 or more profiles will respond if we change the limits, even if they don't actively use all of their profiles. We can assume that a certain amount will upgrade, but a certain amount will also leave Buffer. We also don't quite know what effect word-of-mouth will have. We also may make the assumption that people with only 1 or 2 profiles will be unaffected -- this may not necessarily be the case. We don't fully understand the underlying causal effects that the profile limits have.

One way to get a better understanding of those causal mechanisms would be to run an experiment. If we could adjust the profile limits of a subset of our active Free user population, and allow them to upgrade to increase their limits, we would be able to make a better estimate of how this change might affect our user base and revenue stream. I would suggest running an experiment on a small subset of our active Free user population before making any definitive statement on the effect it will have on revenue. 
