---
date: 2017-09-19T14:56:22-04:00
subtitle: ""
type: "post"
tags: []
title: "How much vacation does Buffer take?"
---

Buffer introduced a [minimum vacation policy](https://open.buffer.com/minimum-vacation/) a year ago, in which we encouraged Americans to take _at least_ 15 days of vacation (excluding holidays), Australians to take at least 17 days, Canadians at least 20 days, and Europeans at least 25 days. 

In this analysis we'll try to measure the effect that it has had on the team over the past year. We use timetastic to schedule time off, and we're lucky enough to have a nice report that shows how much time each team member has taken off.

### Data collection
We'll read the data in from a CSV that we exported from timetastic.

```{r include = FALSE}
library(dplyr); library(buffer); library(ggplot2)
```

```{r}
# read CSV
days <- read.csv('~/Downloads/vacation.csv', header = T)
```

Now let's clean up the column names and set the dates as date types. 

```{r}
# change column names
colnames(days) <- safe_names(colnames(days))

# create function to set date as date object
set_date <- function(column) {
  column <- as.Date(column, format = '%Y-%m-%d')
}

# apply function to date columns
days[c(3, 9:11)] <- lapply(days[c(3, 9:11)], set_date)
```

Great! Now let's do a bit of tidying to make the analysis easier.

### Data tidying
We only want to look at time taken off in the past year, so let's filter out dates that don't apply. We also want to filter to only look at _vacation_ days taken off. We also want to exclude people that are no longer on the team.

```{r include = FALSE}
# people that left
left <- c('Amy Heather', 'Niel de la Rouviere', 'Alex Ray' , 'Patrik Ward', 'Rodolphe Dutel')

# filter out people who left
days <- days %>%
  filter(!(user %in% left))
```


```{r}
# filter dates
vacation <- days %>%
  filter(start_time > (Sys.Date() - 365) & leave_type == "Vacation" & status == "Authorised")
```

Now let's group the data by team member, so that we can see the total number of days taken off by each.

```{r}
# group by person
by_user <- vacation %>%
  group_by(user) %>%
  summarise(total_working_days = sum(working))
```

We'll "bin" the number of vacation days, so that it becomes a categorical variable instead of a numeric one. This will make plotting a little easier, as we'll be able to see the number of team members that took, say, 1 to 5 days in the past year.

```{r}
# make bins for total days
cuts <- c(-Inf, seq(0, 50, 5))

# bin the days
by_user <- by_user %>%
  mutate(days_off = cut(total_working_days, cuts))
```

Alright, let's make some fun plots.

### Exploratory analysis
So how much time has the team taken off in the past year?

```{r echo = FALSE}
by_user %>%
  count(days_off) %>%
  ggplot(aes(x = days_off, y = n)) +
  geom_bar(stat = 'identity') +
  theme_minimal() +
  geom_text(aes(label = n), vjust = -1, color = 'black', size = 4) +
  scale_y_continuous(limits = c(0, 30)) +
  labs(x = "Days Off", y = "Number of Team Members", title = "How much time off did people take?")
```

The average number of vacation days taken by Buffer employees in the past year is 18.2, and the median is 16.5. Around a quarter of the team has taken 12 or less vacation days, and another quarter has taken over 24 days off.

The plot above shows that many team members have only taken 10-15 vacation days. We have one team member who has only taken 2 days off over the past year and one who has taken over 40 days!

How does this look for people that live in different regions? Let's look at the _median_ number of vacation days taken by team members in each region.

```{r include = FALSE}
# give users regions
by_user$region <- c('us', 'asia', 'asia', 'us', 'us', 'europe', 'europe', 'us', 'europe', 'us',
            'us', 'us', 'europe', 'us', 'us', 'us', 'ca', 'europe', 'us', 'europe',
            'europe', 'europe', 'europe', 'us', 'asia', 'europe', 'ca', 'us', 'europe', 'us',
            'europe', 'europe', 'europe', 'us', 'us', 'europe', 'us', 'us', 'europe', 'us',
            'us', 'europe', 'us', 'ca', 'us', 'us', 'europe', 'us', 'europe', 'europe',
            'us', 'europe', 'au', 'europe', 'us', 'asia', 'au', 'europe', 'europe', 'us',
            'ca', 'us', 'asia', 'europe', 'us', 'us', 'us', 'us', 'europe', 'ca', 'us')

# sub region names
by_user$region <- gsub('us', "US", by_user$region)
by_user$region <- gsub('au', "Australia", by_user$region)
by_user$region <- gsub('europe', "Europe", by_user$region)
by_user$region <- gsub('asia', 'Asia', by_user$region)
by_user$region <- gsub('ca', "Canada", by_user$region)
```

```{r echo = FALSE}
by_user %>%
  group_by(region) %>%
  summarise(median = median(total_working_days)) %>%
  mutate(region = reorder(region, median)) %>%
  ggplot(aes(x = region, y = median, fill = region)) +
  geom_bar(stat = 'identity') +
  geom_text(aes(label = round(median, 0)), hjust = 1.5, color = 'white', size = 4) +
  theme_minimal() +
  coord_flip() +
  labs(x = NULL, y = NULL, title = 'Median Vacation Days') +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  theme(text = element_text(size = 15)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(plot.title = element_text(hjust = 0.5)) +
  guides(fill = FALSE) 
```

This is interesting! We haven't yet taken into account the variance in the number of vacation days taken, so let's visualize the distribution of vacation days taken from team members that live in each region.

```{r echo = FALSE, warning = FALSE, message = FALSE}
library(ggridges)

# plot team members and updates
ggplot(filter(by_user, region != 'Australia'), aes(x = total_working_days, 
                                                   y = as.factor(region), 
                                                   fill = region)) +
  geom_density_ridges(rel_min_height = 0.001, scale = 2, color = 'white') +
  theme_ridges() +
  scale_fill_cyclical(values = c("#2A7FFF", "#5599FF")) +
  scale_x_continuous(limits = c(0, 50), breaks = seq(0, 50, 10)) +
  labs(x = "Days Off", y = NULL)
```

Overall, it looks like people are following the recommendations of the policy. The center of the distribution of days off for Americans is between 10 and 20, as is the center of the distribution of days off for Europeans. 

The distribution is far wider for Europeans, which suggests that there are people that have taken very few days off (one person has only taken 2) and people that have taken many days off (one has taken 44.5). 

Canadians seem to know how to relax. Their distribution is more uniformly distributed and is closer to the 30-40 day range. 

Team members in Asia represent a smaller sample, but their distribution is centered around 11 days.

How do the distributions look for people on different teams within Buffer?

```{r include = FALSE}
# read data with teams
by_user <- read.csv('~/Downloads/vaca.csv', header = T)
```

```{r echo = FALSE}
by_user %>%
  group_by(team) %>%
  summarise(median = median(total_working_days)) %>%
  mutate(team = reorder(team, median)) %>%
  ggplot(aes(x = team, y = median, fill = team)) +
  geom_bar(stat = 'identity') +
  geom_text(aes(label = median), hjust = 1.5, color = 'white', size = 4) +
  theme_minimal() +
  coord_flip() +
  labs(x = NULL, y = NULL, title = 'Median Vacation Days') +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  theme(text = element_text(size = 15)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(plot.title = element_text(hjust = 0.5)) +
  guides(fill = FALSE) 
```

This is pretty cool. The executive team is leading the way! In my opinion, this serves as a great example to us underlings to take time to rest and recharge. What do the distributions look like?

```{r echo = FALSE, warning = FALSE, message = FALSE}
ggplot(filter(by_user, team != "Finance"), aes(x = total_working_days, y = team, fill = team)) +
  geom_density_ridges(rel_min_height = 0.001, scale = 1.5, color = 'white') +
  theme_ridges() +
  scale_fill_cyclical(values = c("#2A7FFF", "#5599FF")) +
  scale_x_continuous(limits = c(0, 50), breaks = seq(0, 50, 10)) +
  labs(x = "Days Off", y = NULL)
```

We can see that there is a lot of variance across and within teams. It's really interesting to see the bimodal distributions for the Peopple team and the Execs. Some people within those teams seem to be more willing to take time off for vacation than others. 

### That's it for now
That was fun! Let me know if you have any thoughts or questions!