---
date: 2017-09-19T14:56:22-04:00
subtitle: ""
type: "post"
tags: []
title: "How much vacation does Buffer take?"
---

Buffer introduced a [minimum vacation policy](https://open.buffer.com/minimum-vacation/) a year ago, in which we encouraged Americans to take _at least_ 15 days of vacation (excluding holidays), Australians to take at least 17 days, Canadians at least 20 days, and Europeans at least 25 days. 

The fairness of this policy is open to debate, but we can try to measure the effect that it has had on the team over the past year. We use timetastic to schedule time off, and we're lucky enough to have a nice report that shows how much time each team member has taken off.

### Data collection
We'll read the data in from a CSV that we exported from the site.

```{r include = FALSE}
library(dplyr); library(buffer); library(ggplot2)
```

```{r}
# read CSV
days <- read.csv('~/Downloads/vacation.csv', header = T)
```

Let's clean up the column names and set the dates as date types. 

```{r include = FALSE}
# change column names
colnames(days) <- safe_names(colnames(days))

# create function to set date as date object
set_date <- function(column) {
  column <- as.Date(column, format = '%Y-%m-%d')
}

# apply function to date columns
days[c(3, 9:11)] <- lapply(days[c(3, 9:11)], set_date)
```

Great! Now let's get the data in the format we want.

### Data tidying
We only want to look at time taken off in the past year, so let's filter out dates that don't apply. We also want to filter to only look at _vacation_ days taken off. We also want to exclude people that are no longer on the team.

```{r include = FALSE}
# people that left
left <- c('Amy Heather', 'Niel de la Rouviere', 'Alex Ray' , 'Patrik Ward', 'Rodolphe Dutel')

# filter out people who left
days <- days %>%
  filter(!(user %in% left))
```


```{r}
# filter dates
vacation <- days %>%
  filter(start_time > (Sys.Date() - 365) & leave_type == "Vacation" & status == "Authorised")
```

Now let's group the data by team member, so that we can see the total number of days taken off by each.

```{r}
# group by person
by_user <- vacation %>%
  group_by(user) %>%
  summarise(total_working_days = sum(working))
```

```{r}
# make bins for total days
cuts <- c(-Inf, seq(0, 50, 5))

# bin the days
by_user <- by_user %>%
  mutate(days_off = cut(total_working_days, cuts))
```

## Exploratory analysis
So how much time has the team taken off in the past year?

```{r echo = FALSE}
by_user %>%
  count(days_off) %>%
  ggplot(aes(x = days_off, y = n)) +
  geom_bar(stat = 'identity') +
  theme_minimal() +
  geom_text(aes(label = n), vjust = -1, color = 'black', size = 4) +
  scale_y_continuous(limits = c(0, 30)) +
  labs(x = "Days Off", y = "Number of Team Members", title = "How much time off did people take?")
```

That's interesting! Over the past year, many team members have only taken 10-15 vacation days. A quarter of the team has taken 12 or less days, half of the team has taken 16 or less days, and a quarter of the team has taken 24 or more days. We have one team member who has taken over 40 days of vacation!

How does this look for people that live in different regions? 

```{r include = FALSE}
# give users regions
by_user$region <- c('us', 'asia', 'asia', 'us', 'us', 'europe', 'europe', 'us', 'europe', 'us',
            'us', 'us', 'europe', 'us', 'us', 'us', 'ca', 'europe', 'us', 'europe',
            'europe', 'europe', 'europe', 'us', 'asia', 'europe', 'ca', 'us', 'europe', 'us',
            'europe', 'europe', 'europe', 'us', 'us', 'europe', 'us', 'us', 'europe', 'us',
            'us', 'europe', 'us', 'ca', 'us', 'us', 'europe', 'us', 'europe', 'europe',
            'us', 'europe', 'au', 'europe', 'us', 'asia', 'au', 'europe', 'europe', 'us',
            'ca', 'us', 'asia', 'europe', 'us', 'us', 'us', 'us', 'europe', 'ca', 'us')
```

```{r echo = FALSE, warning = FALSE, message = FALSE}
library(ggridges)

by_user$region <- gsub('us', "US", by_user$region)
by_user$region <- gsub('au', "Australia", by_user$region)
by_user$region <- gsub('europe', "Europe", by_user$region)
by_user$region <- gsub('asia', 'Asia', by_user$region)
by_user$region <- gsub('ca', "Canada", by_user$region)

# plot team members and updates
ggplot(filter(by_user, region != 'Australia'), aes(x = total_working_days, 
                                                   y = as.factor(region), 
                                                   fill = region)) +
  geom_density_ridges(rel_min_height = 0.001, scale = 2, color = 'white') +
  theme_ridges() +
  scale_fill_cyclical(values = c("#2A7FFF", "#5599FF")) +
  scale_x_continuous(limits = c(0, 50), breaks = seq(0, 50, 10)) +
  labs(x = "Days Off", y = NULL)
```

Overall, it looks like people are following the recommendations of the policy! The center of the distribution of days off for Americans is between 10 and 20, as is the center of the distribution of days off for Europeans. 

The distribution is far wider for Europeans, which suggests that there are people that have taken very few days off (one person has only taken 2) and people that have taken many days off (one has taken 44.5). 

Canadians seem to really know how to take time off. Their distribution is more uniformly distributed and is closer to the 30-40 day range. 

Team members in Asia represent a smaller sample, but their distribution is centered around 11 days. Here are the means and medians for each group.

```{r}
by_user %>%
  group_by(region) %>%
  summarise(average = mean(total_working_days), median = median(total_working_days),
            standard_deviation = sd(total_working_days)) %>%
  arrange(average)
```

There we go!
